---
title: "STAT433 Assignment 2"
date: "`r Sys.Date()`"
author: "Yuhan Wang"
output: github_document
---

## load require package

```{r}
library(dplyr)

# install.packages("nycflights13")
library(nycflights13)
library(ggplot2)
library(naniar)
library(lubridate) 
```

# 1) How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
#count the total number of missing dep_time
sum(is.na(flights$dep_time))
```

```{r}
#view the pattern of missing value
vis_miss(flights %>% select(contains("arr_") , contains("dep_")),warn_large_data=FALSE)
```

**We notice that all rows that missing dep_time also missing dep_delay, arr_time and arr_delay, then we infer that they are cancelled flights.**

# 2) Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they are not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
#Convert select columns to minutes since midnight
flights_new <- flights %>% mutate(dep_time_MM = dep_time %/% 100 * 60 + dep_time %% 100,
                                  sched_dep_time_MM = sched_dep_time %/% 100 * 60 + sched_dep_time %% 100)
#Show the result
flights_new %>% select(contains("dep_time"))
```

# 3) Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with \`ggplot(aes(x= ,y=)) + geom_point()'

```{r}
# look at the number of cancelled flights across weekday
flights_new <-  flights_new %>% 
  mutate(departure = make_datetime(year, month, day, hour, minute)) 

flights_new %>% mutate(dayWeek = wday(departure)) %>% group_by(dayWeek) %>% 
  summarise(num_cancel = sum(is.na(dep_time))) %>% 
  ggplot(aes(x=dayWeek,y=num_cancel)) +
  geom_point() + geom_smooth(se = FALSE)
```

**We can see the number of canceled flights is higher for weekdays compared with weekends.**

```{r}
#View(flights %>% filter(is.na(dep_time)))

flights %>% group_by(day) %>% 
  summarise(per_cancel = sum(is.na(dep_time))/n(),
            ave_arr_delay = mean(arr_delay,na.rm = T)) %>% 
  ggplot(aes(x=per_cancel,y=ave_arr_delay)) +
  geom_point() + geom_smooth(se = FALSE)
```

```{r}
flights %>% group_by(day) %>% 
  summarise(per_cancel = sum(is.na(dep_time))/n(),
            ave_dep_delay = mean(dep_delay,na.rm = T)) %>% 
  ggplot(aes(x=per_cancel,y=ave_dep_delay)) +
  geom_point() + geom_smooth(se = FALSE)
```

**We can see the proportion of canceled flights is positively related to the average delay.**
