---
title: "Assignement 3"
author: "Yuhan Wang"
date: "2021/2/27"
output: github_document
---

```{r}
#load require package
library(nycflights13)
library(tidyverse)
library(ggplot2)
library(tidyr)
library(lubridate)
```

Introduction: after summarized the dataset, I have following results:
1. Weekends have less canceled flights and delay departure time compared with weekdays, especially for EWR and LGA airport.
2. Among big airlines(which has more than 10000 recording in the dataset), Delta Airline has the lowest cancel rate and relatively shorter delay time. 
3. Morning have less canceled flights ratio and delay departure time compared with afternoon and evening.


Following three section shows each finding

# 1. choose weekend especially your origin is EWR or LGA

```{r}
# creat week variable
flights_new <-  flights %>% 
  mutate(departure = make_datetime(year, month, day, hour, minute),
         dayWeek = wday(departure),
         dayWeek = ifelse(dayWeek==1,'Sun',ifelse(dayWeek==2,'Mon',ifelse(dayWeek==3,'Tue',
                   ifelse(dayWeek==4,'Wed',ifelse(dayWeek==5,'Thu',ifelse(dayWeek==6,'Fri',
                   ifelse(dayWeek==7,'Sat','Missing')))))))) 

flights_new$dayWeek <- factor(flights_new$dayWeek, levels = c('Mon','Tue','Wed','Thu','Fri','Sat','Sun'))
flights_new$origin <- factor(flights_new$origin, levels = c('LGA','EWR','JFK'))

# look at the number of canceled flights across weekday
flights_new %>% group_by(dayWeek) %>%
  summarise(num_cancel = sum(is.na(dep_delay)),
            per_cancel = num_cancel/n(),
            ave_delay = mean(dep_delay,na.rm=T))
```

```{r}
#plot the number of cancelled flights by weekday and origin airports
flights_new %>% 
  filter(is.na(dep_delay)) %>% 
  ggplot(aes(x=dayWeek,fill = origin)) +
  geom_bar(position=position_dodge())
```

# 2. choose Delta Airline if possible

```{r}
table(flights$carrier)
```


```{r}
# take airline that has more than 10000 flights records
big_airline_list <- c('9E','AA','B6','DL','EV','MQ','UA','US','WN')

flights2 <- flights %>%
  left_join(airlines, by = "carrier")

# look at the number of canceled flights across airlines
flights2 %>% filter(carrier %in% big_airline_list) %>% group_by(name) %>%
  summarise(num_cancel = sum(is.na(dep_delay)),
            per_cancel = num_cancel/n(),
            ave_delay = IQR(dep_delay,na.rm=T)) %>% 
  arrange(per_cancel)
```

```{r,warning=FALSE}
# make a density plot dep_time across big airlines

flights_big <- flights2 %>% 
  filter(carrier %in% big_airline_list) 

flights_big %>%
 filter(dep_delay >= -43L & dep_delay <= 119L | is.na(dep_delay)) %>%
 ggplot() +
 aes(x = dep_delay, fill = name) +
 geom_density(adjust = 1L) +
 scale_fill_hue() +
 theme_minimal() +
 facet_wrap(vars(name))
```






# 3. choose morning to take your flights

```{r}
# creat time period variable
flights3 <- flights %>% 
  mutate(time_period = ifelse(hour<=12,'morning',
                       ifelse(hour>12 & hour<19,'afternoon',
                       ifelse(hour>=19,'evening','missing'))))

flights3$time_period <- factor(flights3$time_period, levels = c('morning','afternoon','evening'))

# look at the number of canceled flights across time periods
flights3 %>% group_by(time_period) %>%
  summarise(num_cancel = sum(is.na(dep_delay)),
            per_cancel = num_cancel/n(),
            ave_delay = mean(dep_delay,na.rm=T)) %>% 
  arrange(per_cancel)
```


```{r,warning=FALSE}
# show the violin polt of delay time across time period
flights3 %>% 
  filter(!is.na(dep_delay)) %>% 
  ggplot(aes(x=time_period,y = dep_delay, fill = time_period)) +
  geom_violin() + ylim(-30,100)
```

