---
title: "In-class exercise"
author: "Yuhan Wang"
date: "2021/2/26"
output: github_document
---

## load require package

```{r}
library(dplyr)
library(tidyr)

# install.packages("nycflights13")
library(nycflights13)
library(ggplot2)
library(lubridate) 
```

# 1. Make a table that describes each plane. It should have a column for tailnum, another column for average arrival delay, and another for the year the plane was manufactured.

```{r}
flights %>% 
  group_by(tailnum) %>% 
  summarise(mean_arr_delay = mean(arr_delay, na.rm = T)) %>% 
  left_join(planes) %>% 
  select(tailnum:year)
```


# 2. Make a table where each row is a day of the year. The first column is the date. The 2:4 columns give the number of (scheduled) departures from EWR, LGA, and JFK.

```{r}
flights %>% 
  mutate(date = as.Date(time_hour)) %>% 
  group_by(origin,date) %>% 
  summarise(deps = n()) %>% 
  pivot_wider(names_from = origin, values_from = deps)
```

# 3. Make a table where each row is a day of the year. Each destination airport is a column. The elements (day x destination) give the number of flights to that destination. What should NA’s be?

```{r}
flights %>% 
  mutate(date = as.Date(time_hour)) %>% 
  group_by(dest,date) %>% 
  summarise(deps = n()) %>% 
  pivot_wider(names_from = dest, values_from = deps)
```




